Building socket-aware BPF programs

### 2018-11-01 paper/paper.tex
paper: Address feedback from Jarno

    To build such awareness into a BPF helper, the verifier needs the ability
    to track the safety of the calls, including appropriate reference counting
    upon the underlying socket. This paper describes extensions to the verifier
    to perform tracking of references in a BPF program. This allows BPF
    developers to extend the UAPI with functions that allocate and release
    resources within the execution lifetime of a BPF program, and the verifier
    will validate that the resources are released exactly once prior to program
    completion.

@@

In recent years, multiple efforts have been made to build new filtering
implementations based upon
BPF~\cite{cilium,ovs-ebpf,nsx-harmony,droplet,xdp-ddos,alfwrapper}. Many of
these implementations aim to provide the functionality of a stateful
firewall~\cite{stateful-fw} - that is, to allow network policies to be defined
to allow traffic from one network endpoint to talk to another network endpoint,
and to implicitly allow responses to pass back through the firewall. Many
implementations over the past decade have developed this functionality in a
manner that is agnostic to the locality of the endpoints. This paper revisits
the possibilities for stateful firewalling when the endpoint socket is
co-located with the kernel providing firewalling functionality, effectively
combining the ideas of \cite{ipfw,stateful-fw,dfw}.

@@

With the verifier now aware of socket pointer types, this could be built upon
to implicitly associate socket pointers with references to kernel resources.

@@

The verifier is tasked with ensuring that all BPF program execution flow is
safe to the extents where it interacts with the BPF API. In general BPF
programs---once verified and triggered by event hooks---will execute from the
first instruction through the final instruction for a particular path through
the program. Therefore, references to those resources should not leak if the
program has been verified.

@@

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this is
intuitive. The information that is needed to perform stateful firewalling is
not present on the Linux instance that needs this information, so to understand
the state of the peers the information must be ether shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be synthesised based upon the packets that are seen on
this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one peer is co-located with the Linux stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

@@

A common design of devices providing routing and forwarding functionality is to
place high-bandwidth forwarding elements into a system with a low-bandwidth
CPU. In many cases, the goal of the CPU is to only to handle control traffic
for configuring the forwarding path, which is a fraction of the traffic handled
by the actual forwarding path. Existing implementations may direct only traffic
for a specific IP into the management CPU, or could plausibly use XDP to
configure DoS filters so that only legitimate traffic is forwarded up the
stack. Such implementations may not provide the degree of restriction on
locally destined traffic that is desired, or may require additional
co-ordination with the socket layer that requires significant orchestration.

@@

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann,
Martin KaFai Lau, Nitin Hande and Thomas Graf for guidance, review and testing
of the patches.

>>>>

    To build socket introspection into a BPF helper, the verifier needs the
    ability to track the safety of the calls, including appropriate reference
    counting upon the underlying socket. This paper describes extensions to the
    verifier to perform tracking of references in a BPF program. This allows
    BPF developers to extend the UAPI with functions that allocate and release
    resources within the execution lifetime of a BPF program, and the verifier
    will validate that the resources are released exactly once prior to program
    completion.

@@

In recent years, multiple efforts have been made to build new filtering
implementations based upon
BPF~\cite{cilium,ovs-ebpf,nsx-harmony,droplet,xdp-ddos,alfwrapper}. Many of
these implementations aim to provide the functionality of a stateful
firewall~\cite{stateful-fw} - that is, network policies may be defined to allow
traffic from one network endpoint to talk to another network endpoint, and to
implicitly allow responses to pass back through the firewall. Many
implementations over the past decade have developed this functionality in a
manner that is agnostic to the locality of the endpoints. This paper revisits
the possibilities for stateful firewalling when the endpoint socket is
co-located with the kernel providing firewalling functionality, as is the case
with container workloads. This effectively combines the ideas of a BPF-based
firewall~\cite{ipfw} with the model of a stateful firewall~\cite{stateful-fw},
operating with an understanding of socket state~\cite{dfw}.

@@

With the verifier now aware of socket pointer types, this provides a foundation
to associate socket pointers with references to kernel resources.

@@

The verifier is tasked with ensuring that all BPF program execution flow is
safe to the extent where it interacts with the BPF API. In general BPF
programs---once verified and triggered by event hooks---will execute from the
first instruction through the final instruction for a particular path through
the program. Therefore, references to those resources should not leak if the
program has been verified.

@@

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this is
intuitive. The information that is needed to perform stateful firewalling is
not present on the Linux instance that needs this information, so to understand
the state of the peers the information must be either shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be reconstructed based upon the packets that are seen
on this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one peer is co-located with the Linux stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

@@

A common design of devices providing routing and forwarding functionality is to
place high-bandwidth forwarding elements into a system with a low-bandwidth
CPU. In many cases, the goal of the CPU is to only to handle control traffic
for configuring the forwarding path, which is a fraction of the traffic handled
by the actual forwarding path. Existing implementations may direct only traffic
for a specific IP into the management CPU, or could redirect all traffic to the
management CPU and configure DoS filters in XDP so that only legitimate traffic
is forwarded up the stack. Such implementations may not provide the degree of
restriction on locally destined traffic that is desired, or may require
additional co-ordination with the socket layer that requires significant
orchestration.

@@

The author would like to acknowledge Alexei Starovoitov, Daniel Borkmann, Jarno
Rajahalme, Martin KaFai Lau, Nitin Hande and Thomas Graf for guidance, review
and testing of the patches, and feedback on this paper.
